NMRLipids VI: Polarizable Force Fields

### 2023-08-11 manuscriptPOL.tex
Updates from Overleaf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Conventional classical molecular dynamics (MD) simulations only consider electrostatics as interactions between static point charges, set by molecular dynamics models (force fields) as an input in the beginning of the simulation. The dynamic effects arising from polarizability are ignored or at most enter the simulations in an averaged fashion in the force field parametrization process. However, since decades work~\cite{Thole1981,ando2001stable,Grossfield2003,
lamoreux2003,Antila2013 ,Lemkul2016, baker2015polarizable,jing2019polarizable } has been underway to make MD simulations explicitly polarizable in the hopes of reaching a more accurate description of the system.

@@

Historically, there has been three main approaches to modelling polarizability: 1) the induced point dipole/multipole model where polarizable point dipoles are placed on chosen sites on a molecule, 2) the classical Drude oscillator model where polarization is modelled by separating a core and a shell charge with a spring which stretches according to the enviroment giving the site a dipole moment, and 3) the electronegativity equilization (fluctuating charge) model where atomic charges are no longer constant but can change according to the electronegativities of the molecule atoms and the electric fields from the molecular enviroment. In recent years the first two have surpassed the third in popularity. However, all of these approaches result in an increasing computational cost, e.g, by introducing new types of interactions, more interactions sites, or by requiring a shorter time step in the simulations.

@@

Considering the increasing popularity of both implicitly and explicitly polarizable force fields and the increased computational cost that they require, a systematic benchmark study on polarizable force field quality is in great need. The few published studies reviewing the state of polarizable force fields for biomolecule simulations did not go beyond citing results from the existing studies~\cite{inakollu2020polarisable,jing2019polarizable,baker2015polarizable}. Recent publications show that \textit{implicit} inclusion of polarizability improves cation binding simulations when evaluated against NMR data~\cite{Melcr:2018a, melcr2019improved}, which suggest that including polarization explicitly could result in better description of ion binding as well. 
Currently, there are three main polarizable lipid force fields with explicit electronic polarization: CHARMM-Drude~\cite{li2017drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}. 
Here we evaluate the performance of two actively developed polarizable force fields with lipid parameters available, CHARMM-Drude~\cite{li2017drude} and AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, against NMR data. We use NMR order parameters to evaluate  structural properties and cation binding, whereas NMR relaxation rates and correlation times are utilized to benchmark the lipid conformational dynamics. Small angle scattering (SAXS) data is used to estimate the description of electron density profile of the membrane. Furthermore, we compare the results to the best performing non-polarizable simulations from the NMRlipids Databank\cite{Databank}.

@@

\section{Results}

@@

One can clearly see that the non-polarizable force fields, CHARMM36 and OPLS3e, outperform the polarizable force fields with respect to the NMR measurables for all parts of the molecule (higher $P$-values). For CHARMM-Drude force field, this is likely due to the headgroup and glycerol parametrization having been done targetting absolute average values of the order parameters instead of taking account the potential order parameter forking and order parameter sign~\cite{Antila2022}. The Amoeba force field reproduces the headgroup order parameters well [\todo{HANNE: refer to the hg flexibility? @Samuli, I didn't quite understand the argument? Refer there to the dihedral figure}] but clearly fails to depict the tail conformations, in particular in the double bond region (Fig.~\ref{fig:order_parameters}). In the tail regions, the CHARMM-Drude model is clearly better out of the two polarizable force fields.

The higher $FFq$ values (Table~\ref{tab:my_label}) of the polarizable force fields indicate that they are not able to reproduce the SAXS data as well as the best non-polarizable force fields. That said, CHARMM-Drude is as good as (POPC) or better (POPE) than its non-polarizable counterpart CHARMM36. AMOEBA, on the other hand, captures the SAXS profile of POPE very poorly (Table~\ref{tab:my_label} and Fig.~\ref{fig:order_parameters}D top). Note that the method for estimating the charge density of bilayers from the simulations places electrons as point charges at the atom centers in the system and does not consider any redistribution of charge density due to the polarizability. We expect this to results into a good estimate of the charge density.

We proceed to quantify (Fig.~\ref{fig:correlation_times}) the structural dynamics of the polarizable models againts NMR $R_{1}$ rates, measuring a window of dynamics around $\sim$0.1-1~ns, and effective correlation times $\tau_{\mathrm{eff}}$ which quantifies a larger range of dynamical processes up to $\sim$1000~ns. We include the same measurables quantified from non-polarizable CHARMM36 and SLipids force fields, which are the best MD force fields regarding the dynamics~\cite{Antila2021}, as a point of comparison. The Amoeba force field captures the dynamics with respect to both measurables well,  in par with the best non-polarizable models. In contrast, the dynamics of CHARMM-Drude appears to be faulty in the fast range ($R_{1}$ rates) and, furthermore, $\tau_{\mathrm{eff}}$ values reveal vastly too slow overall dynamics. This indicates that not only are the simulations with this polarizable force field computationally costly but one would have to simulate longer to obtain converged results. Note that the experimental data here is from POPC. However, chemically identical the headgroup region the dynamics is assumed to be very similar in DOPC, due to the decoupling of the dynamics between the headgroup and the glycerol regions~\cite{Antila2022rot, Klauda08}.

Finally, we address the premise that adding polarizability to force fields improves the depiction of ion binding to biomembranes in simulations. For this, we use the concept of "lipid electrometer"~\cite{Catte2016} where the amount of ion binding to the membrane can be quantified by the change in $\alpha$ and $\beta$ order parameters upon increasing salt concentration. These changes in order parameters are depicted in Fig.~\ref{fig:popc_order_parameter_change} whereas Fig.~\ref{fig:ion_density_profiles} gives the density profiles of ions with respect to the bilayer normal. The CHARMM36-Drude appears to have strong sodium binding, evidenced both by large change in the order parameters and the ion density profiles. The binding of monovalent sodium is, in fact, stronger than the divalent calcium. In strong contrast, experimentally one sees no sodium binding and some binding of calcium ions.

As opposed to the CHARMM36-Drude model, the order parameters of the AMOEBA model are almost completely unaffected by the ion addition, both for NaCl and CaCl$_{2}$. The ion density profiles confirm that very little ions penetrate the membrane region where the lipid phosphates reside. While this is in line with experiments regarding sodium ions, there should be some binding of calcium ions to the membrane. Again, the non-polarizable CHARMM36 appears to depict this calcium binding the best out of the models compared here.

%\begin{figure}[!hbt]
%	\centering
%	\includegraphics{popc_order_parameters_nosalt.png}
%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPC simulations without salt.}
%	\label{fig:popc_order_no_salt}
%        \todo{SAMULI: It would be good to put the experimental value also to second hydrogen even though it would be the same as for first.
%          Now it looks like that those values would not be known.} \\
%        \todo{SAMULI: y-axis should be more zoomed in.} \\
%        \todo{SAMULI: Similar figure for POPE (now figure 4) could be here as panel B).}
%\end{figure}

%\begin{figure}[!hbt]
%	\centering
%	\includegraphics{pope_order_parameters.png}
%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPE simulations.}
%	\label{fig:pope_order}
%        \todo{SAMULI: Move to panel B) in Figure 1.} \\
%        \todo{SAMULI: y-axis should be more zoomed in.} 
%\end{figure}

\begin{figure}[!hbt]
    \centering
    %\includegraphics{Figures/order_parameters.png}
    \includegraphics[width=\textwidth]{Figures/quality.pdf}
    \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank.}
    \label{fig:order_parameters}

@@

\begin{table}[]
    \centering
    \begin{tabular}{c c c c c c c}
        Lipid & Force field & $P^{\mathrm{headgroup}}$ & $P^{sn-1}$ & $P^{sn-2}$ & $FF_{q}$ & APL \\
        \hline
        POPC & OPLS3e       & 0.76 & 0.87 & 0.85 & 0.15 & 66.5\\
        POPC & CHARMM36     & 0.70 & 0.54 & 0.69 & 1.16 & 65.0\\
        POPC & CHARMM-Drude & 0.51 & 0.22 & 0.47 & 1.16 & 62.5\\
        \hline
        POPE & GROMOS-CKP & 0.29 & 0.83 & 0.48 & 0.40 & 59.6\\
        POPE & CHARMM36   & 0.54 & 0.52 & 0.27 & 1.30 & 57.2 \\
        POPE & CHARMM-Drude & 0.06 & 0.53 & 0.27 & 0.80 & 56.6  \\
        POPE & AMOEBA & 0.21 & 0.10 & 0.23 & 3.80 & 66.9\\
        \hline
        DOPC & AMOEBA & 0.60 & 0.60 & 0.54 & - & 70.2\\
    \end{tabular}
    \caption{NMRlipids databank quality evaluations~\cite{Databank} and area per lipids compared between the best simulations found from the NMRlipids databank, simulations with CHARMM36 force field, and simulations with polarizable force fields. Experimental estimations for areas per lipid are POPC: 64.3$\pm1$\AA$^{2}$\cite{kucerka2011}, DOPC: 67.5$\pm1$\AA$^{2}$\cite{kucerka2008}, and POPE: 56.7$\pm$3\AA$^{2}$\cite{Rickeard2020}}
    \label{tab:my_label}
\end{table}

@@

\begin{figure}[!hbt]
	\centering
	\includegraphics{Figures/both_log.pdf}
	\caption{Effective correlation times (top) and R$_1$ rates (bottom) for the polarizable models and best performing non-polarizable force fields. Note, top panel axis is logarithmic to visualize the slow CHARMM-drude dynamics. Experimental values are obtained from Ref.~\citenum{Antila2022rot}. All simulations used here were salt free.}
	\label{fig:correlation_times}

@@

%\begin{figure}[!hbt]
%	\centering
%	\includegraphics{popc_order_parameters_for_all.png}
%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPC simulations. The experimental values on the dashed lines mean that the different isomers 1/2 experimentally could not be resolved.}
%	\label{fig:popc_order}
%        \todo{SAMULI: I am not sure if we need this kind of figure. Systems without ions are in separate figures as well as dependence on ions.}
%\end{figure}

@@

\begin{figure}[!hbt]
    \centering
    \includegraphics{Figures/ion_density_profiles_with_chloride.eps}
    \caption{Density profiles along the membrane normal for the Drude 2017 and 2023 POPC (first and second rows) and AMOEBA DOPC (third row). Solid and dashed lines are for the systems with the NaCl and CaCl$_{2}$, respectively.}
    \label{fig:ion_density_profiles}
            \todo{SAMULI: We should also show the ion density profiles along membrane normal and compare the ones with 1000 mM NaCl and 350 mM CaCl$_2$
          (concentration in bulk water) to the best profiles in the literature (Fig. 5 in http://dx.doi.org/10.1021/acs.jpcb.7b12510).
          This will show the actual difference in ion binding to the best performing models available.}
          \todo{Batuhan: Data generated from http://dx.doi.org/10.1021/acs.jpcb.7b12510}
\end{figure}

@@

%\subsection{Discussion}

\subsection{Conclusions}
After comparing the polarizable lipid models against high-resolution experimental data and the best performing non-polarizable force fields, we cannot confirm that adding polarizablity to the force fields has yet solved many of the issues it was hoped to address. Ion binding to the model membrane seems to be overestimated by the CHARMM36-Drude model, whereas the calcium binding is underestimated by the Amoeba force field. The electron density profiles of the bilayers appear to be better reproduced by the best non-polarizable force fields, within the confines of the approximations used to calculate them from the trajectory. Same applies for the dynamics and conformational sampling of the models, although the Amoeba force field produces results, spesifically in the headgroup region of PC lipids, that are in par or better than the best non-polarizable force fields with respect to these measurables. In summary, polarizable models appear still to be in their infancy, and added complexity of polarizability has not still realised its potential to the fullest; The non-polarizable force fields with longer and more intense refinement behind them still outpeform. Considering the complexity and cost of using these models, polarizability should be parameterized with more thorough consideration to the experimental benchmarks.

@@

In the Amoeba force field, a set of atomic multipoles were first obtained based on the QM electron densities using Stone's distributed multipole analysis~\cite{Stone1981}. The atoms were then assigned polarizabilities according to the AMOEBA atom types and induced dipole contribution was self consistently subtracted from the permanent dipoles~\cite{shi2013proteinamoeba}. The charges (monopoles) were then kept constant while the multipoles were refined against the electrostatic potential energy surface around few conformers of the molecular fragments. %Multipoles are allowed to induce only dipoles at polarizable sites outside their immediate environment (polarization group). 
After electrostatic parameters were set, the Lennard-Jones parameters were taken from previous AMOEBA parameterizations of smaller molecules containing same functional groups~\cite{Ren2011polorganic,shi2011hydration}, and then refined against molecule - water (gas phase) interaction energies. After settling all the nonbonded parameters, dihedral potentials were parameterised based on the QM conformational energy profiles.

@@

Notes on Amoeba parametrization (this is NOT well explained in the paper)
-bottom up approach
-Stone distributed multipole analysis to get the multipoles from the QM electron densities. Divided into permanent and induced.
-Then iterative fitting to QM-PES for electrostatics
-WdW from 1) optimization to hydration energy calculations and 2) QM fragment-water gas phase calculations.
-dihedrals optimized to QM potentials
-no mention of the screening for polarization? Do they use some AMOEBA value
\end{comment}
\subsubsection{Using a polarizable force field}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\textcolor{red}{some ideas for the intro/conclusion: 1) amoeba and drude are mostly used for transmembrane protein simulations. https://pubs.acs.org/doi/abs/10.1021/acsnano.1c06443 https://www.sciencedirect.com/science/article/abs/pii/S0959440X2100110X https://pubs.rsc.org/en/content/articlehtml/2021/sc/d1sc01887f etc. Validation/comparison of these simulations with the experiments are done for the proteins, not the lipids. Now since we know that the lipid dynamics and conformations are not very correct with the polarizable force fields, can these affect the protein conformations? is there a way to possibly connect this to the previous nmrlipids jacs paper?\\
}

Conventional classical molecular dynamics (MD) simulations only consider electrostatics as interactions between static point charges, set by molecular dynamics models (force fields) as an input in the beginning of the simulation. The dynamic effects arising from polarizability are ignored or at most enter the simulations in an averaged fashion in the force field parametrization process. However, since decades work~\cite{Thole1981,ando2001stable,Grossfield2003,
lamoreux2003,Antila2013 ,Lemkul2016, baker2015polarizable,jing2019polarizable } has been underway to make MD simulations explicitly polarizable in the hopes of reaching a more accurate description of the system.

@@

Historically, there has been three main approaches to modelling polarizability: 1) the induced point dipole/multipole model where polarizable point dipoles are placed on chosen sites on a molecule, 2) the classical Drude oscillator model where polarization is modelled by separating a core and a shell charge with a spring which stretches according to the enviroment giving the site a dipole moment, and 3) the electronegativity equilization (fluctuating charge) model where atomic charges are no longer constant but can change according to the electronegativities of the molecule atoms and the electric fields from the molecular enviroment. In lipid modelling, the first two have surpassed the third in popularity. All of these approaches result in an increasing computational cost, e.g, by introducing new types of interactions, more interactions sites, or by requiring a shorter time step in the simulations.

@@

Considering the growing popularity of both implicitly and explicitly polarizable force fields and the increased computational cost that they require, a systematic benchmark study on polarizable force field quality is in great need. The few published studies reviewing the state of polarizable force fields for biomolecule simulations did not go beyond citing results from the existing studies~\cite{inakollu2020polarisable,jing2019polarizable,baker2015polarizable}. Recent publications show that \textit{implicit} inclusion of polarizability improves cation binding simulations when evaluated against NMR data~\cite{Melcr:2018a, melcr2019improved}, which suggest that including polarization explicitly could result in better description of ion binding as well. 
Currently, there are three main polarizable lipid force fields with explicit electronic polarization: CHARMM-Drude~\cite{li2017drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}. 
Here we evaluate the performance of two actively developed polarizable force fields with lipid parameters available, CHARMM-Drude~\cite{li2017drude} and AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, against NMR data. We use NMR order parameters to evaluate  structural properties and cation binding, whereas NMR relaxation rates and correlation times are utilized to benchmark the lipid conformational dynamics. Small angle scattering (SAXS) data is used to estimate the description of electron density profile of the membrane. Furthermore, we compare the results to the best performing non-polarizable simulations from the NMRlipids Databank\cite{Databank}.

@@

\section{Results and Discussion}

@@

One can clearly see that the non-polarizable force fields, CHARMM36 and OPLS3e, consistently outperform the polarizable force fields with respect to the NMR measurables for all parts of the molecule (higher $P$-values). For older CHARMM-Drude force field, this is likely due to the headgroup and glycerol parametrization having been done targetting absolute average values of the order parameters instead of taking account the potential order parameter forking ( measurably different order parameters for different C--H bods at one carbon) and order parameter sign~\cite{Antila2022}. This was then remedied in the newer version of the model, which also provides better description of the carbon tails. In fact, for POPE sn-2 tail the CHARMM-Drude2023 gives the best description of the models investigated here.  The AMOEBA force field reproduces the most of the headgroup order parameters well for both DOPC and POPE but vastly overestimates one of the g$_3$ values, resulting in too large forking of oder parameters in this particular carbon. AMOEBA also clearly struggles to depict the tail conformations, in particular in the double bond region of DOPC (Fig.~\ref{fig:order_parameters}). In the tail regions, the CHARMM-Drude2023 appears to be the best out of the polarizable models.

The higher $FFq$ values (Table~\ref{tab:my_label}) of the polarizable force fields for POPC indicate that they are not able to reproduce the SAXS data quite as well as the best non-polarizable force field, OPLS3e. That said, CHARMM-Drude2023, in particular, is clearly better than its non-polarizable counterpart CHARMM36 and the best out of all the models here in capturing the POPE SAXS data. AMOEBA, on the other hand, captures the SAXS profile of POPE very poorly (Table~\ref{tab:my_label} and Fig.~\ref{fig:order_parameters}D top) which is likely connected to the overall description of the lipid tails in the force fields. Note that owing the simulation size dependency of the SAXS data~\cite{Databank} one should focus on the extrema $q_{z}$[\AA$^{-1}$] when comparing the simulated and experimental data. In particular $FFq$ only measures the difference between the first minima of the simulated and experimental curves. Also note, that the method for estimating the charge density of bilayers from the simulations places electrons as point charges at the atom centers in the system and does not consider any redistribution of charge density due to the polarizability. Nevertheless, we expect this results into a good estimate of the charge density for the calculation of the SAXS curve.

\subsection{Evaluation of conformational dynamics of lipids}

We proceed to evaluate (Fig.~\ref{fig:correlation_times}) the structural dynamics of the polarizable models againts NMR $R_{1}$ rates, here measured at a magnetic frequency that captures a window of dynamics around $\sim$0.1-1~ns, and effective correlation times $\tau_{\mathrm{eff}}$ which quantifies a larger range of dynamical processes up to $\sim$1000~ns. We include the same measurables calculated from non-polarizable CHARMM36 and SLipids force fields---the best MD force fields regarding the dynamics~\cite{Antila2021}---as a point of comparison. The AMOEBA force field captures the dynamics with respect to both measurables very well,  in par with the best non-polarizable models. In contrast, the dynamics of CHARMM-Drude appears to be faulty in the fast range ($R_{1}$ rates) and, furthermore, $\tau_{\mathrm{eff}}$ values reveal vastly too slow overall dynamics. The updated CHARMM-Drude2023 does not significantly improve the dynamics at the fast range investigated here. The effective correlation times of the newer Drude model are improved but still an order of magnitude too slow. This indicates that not only are the simulations with this polarizable force field computationally costly but one would have to simulate longer to obtain converged results. The non-polarizable counterpart of the Drude model, CHARMM36, exhibits much more realistic, faster effective correlation times. Interestingly, it has also been observed that CHARMM-Drude models have slower dynamic of water hydrogen bonding around amino acids compared to their non-polarizable counterpart~\cite{Ngo2019} which might align with an overall slower dynamics of the model in addition to enhanced water binding. Note that the experimental data here is from POPC. However, chemically identical the headgroup region the dynamics is assumed to be practically identical in DOPC, due to the decoupling of the dynamics between the headgroup and the glycerol regions~\cite{Antila2022rot, Klauda08}.

\subsection{Cation binding to membranes in polarizable simulations}

Finally, we address the premise that adding polarizability to force fields improves the depiction of ion binding to biomembranes in simulations. For this, we use the concept of "lipid electrometer"~\cite{Catte2016} where the amount of ion binding to the membrane can be quantified by the change in $\alpha$ and $\beta$ order parameters upon increasing salt concentration. These changes in order parameters are depicted in Fig.~\ref{fig:popc_order_parameter_change} whereas Fig.~\ref{fig:ion_density_profiles} gives the density profiles of ions with respect to the bilayer normal. The CHARMM-Drude2017 exhibits strong sodium binding, evidenced both by large change in the order parameters and the ion density profiles. Based on the change in the order parameters, the sodium binding is even stronger than the divalent calcium binding. In strong contrast, experimentally one sees no sodium binding and some binding of calcium ions. Too strong sodium binding in the CHARMM-Drude model has been indicated before for peptides/amino acids~\cite{Ngo2019, Kav2022} and deep-eutectic solvents~\cite{shayestehpour2022ion} along with a too strong calcium interactions~\cite{Tan2022} in line with what we observe here for the lipids.

The faulty sodium ion binding seems to have been somewhat remedied in the CHARMM-Drude2023 force field: the change of the order parameters is diminished to be more in line with experiments and only the some accumulation of sodium ions near the phosphate oxygen is seen in the ion distributions. Based on the change in the order parameters, the calcium binding also appears to be better represented in the updated model. The ion distributions reveal that binding of calcium has moved deeper into the bilayer and there is a stronger formation of ionic double layer by subsequent attraction of chloride into the bilayer. However, this the ion layering is most prominently present in the original CHARMM36 force field, for for sodium and calcium ions. Note, that both for sodium and calcium ion additionin the CHARMM-Drude2023force field reintroduces the forking of order parameters at the $\beta$ carbon in particular. These might also be a sign of poor convergence of the simulations owing to the slow conformational dynamics detected above, which would also not be fully reflected in the calculation of error estimates for the order parameters.

As opposed to the CHARMM36-Drude model, the order parameters of the AMOEBA model are almost completely unaffected by the ion addition, both for NaCl and CaCl$_{2}$. The ion density profiles confirm that very little ions penetrate the membrane region where the lipid phosphates reside. While this is in line with experiments regarding sodium ions, there should be some binding of calcium ions to the membrane. Again, the change in the order parameters indicate the non-polarizable CHARMM36 depicts sodium binding the best out of the models compared here, being slightly more realistic than its most recent polarizable counterpart, CHARMM-Drude2023. Regarding calcium binding, CHARMM36 captures the $\Delta^{\alpha}_{CH}$ about as well as the CHARMM-Drude2023, but the change in the $\beta$ carbon is better represented in the CHARMM-Drude2023 and AMOEBA models. This suggest that some accuracy might have been gained in modelling the divalent ion binding with polarizable models but the advantage is not clear. No force field, polarizable or non-polarizable, is able to model both monovalent and divalent ion binding flawlessly. This is of note, as polarizable force fields are often assumed superior and concequently a popupar choice in modelling, eg, ion channels~\cite{sun2017 jing2019polarizable, klesse2020induced}\todo{I think there is a mistake in the C36 data, firsly, the concentration should probably be 950mM instead of 1000mM and secondly the monovalent anion and cation curves look identical} ~\todo{batuhan: I fxed the anion and cation curves. Yes, C36 data is at 950mM but explicitly adding this as a new color could clutter the figure. Should I just mention this in the caption?}.

\begin{figure}[t]
    \centering
    %\includegraphics{Figures/order_parameters.png}
    \includegraphics[width=0.4\textwidth]{Figures/molecules.pdf}
    \caption{Molecules and atom labelling used.}
    \label{fig:molecules}

@@

\begin{figure}[!hbt]
    \centering
    %\includegraphics{Figures/order_parameters.png}
    \includegraphics[width=0.8\textwidth]{Figures/quality.pdf}
    \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank. The atom labelling is as depicted in Fig.~\ref{fig:molecules}.}
    \label{fig:order_parameters}

\begin{table}[]
    \centering
    \begin{tabular}{c c c c c c c}
        Lipid & Force field & $P^{\mathrm{headgroup}}$ & $P^{sn-1}$ & $P^{sn-2}$ & $FF_{q}$ & APL \\
        \hline
        POPC & OPLS3e       & 0.76 & 0.87 & 0.85 & 0.15 & 66.5\\
        POPC & CHARMM36     & 0.70 & 0.54 & 0.69 & 1.16 & 65.0\\
        POPC & CHARMM-Drude & 0.51 & 0.22 & 0.47 & 1.16 & 62.5\\
        POPC & CHARMM-Drude2023 & 0.50 & 0.49 & 0.53 & 0.82 & 64.5\\
        \hline
        POPE & GROMOS-CKP & 0.29 & 0.83 & 0.48 & 0.40 & 59.6\\
        POPE & CHARMM36   & 0.54 & 0.52 & 0.27 & 1.30 & 57.2 \\
        POPE & CHARMM-Drude & 0.06 & 0.53 & 0.27 & 0.80 & 56.6  \\
        POPE & CHARMM-Drude2023 & 0.28 & 0.59 & 0.54 & 0.00 & 61.4  \\
        POPE & AMOEBA & 0.21 & 0.10 & 0.23 & 3.80 & 66.9\\
        \hline
        DOPC & AMOEBA & 0.60 & 0.60 & 0.54 & - & 70.2\\
    \end{tabular}
    \caption{NMRlipids databank quality evaluations~\cite{Databank} and area per lipids compared between the best simulations found from the NMRlipids databank, simulations with CHARMM36 force field, and simulations with polarizable force fields. Experimental estimations for areas per lipid are POPC: 64.3$\pm1$\AA$^{2}$\cite{kucerka2011}, DOPC: 67.5$\pm1$\AA$^{2}$\cite{kucerka2008}, and POPE: 56.7$\pm$3\AA$^{2}$\cite{Rickeard2020}.}
    \label{tab:my_label}
\end{table}

@@

\begin{figure}[!hbt]
	\centering
	\includegraphics{Figures/both_inkscp.pdf}
	\caption{Effective correlation times (top) and R$_1$ rates (bottom) for the polarizable models and best performing non-polarizable force fields. Note, top panel axis is logarithmic to visualize the slow Drude dynamics. Experimental values are obtained from Ref.~\citenum{Antila2022rot}. All simulations used here were salt free.}
	\label{fig:correlation_times}

@@

@@

\begin{figure}[!hbt]
    \centering
    \includegraphics{Figures/ion_density_profiles_with_chloride.eps}
    \caption{Density profiles along the membrane normal for the CHARMM36 (first row)~\cite{Catte2016}, Drude 2017 and 2023 POPC (second and third rows), AMOEBA DOPC (fourth row), and electronic continuum correction (ECC, last row)~\cite{Melcr:2018a}. Solid and dashed lines are for the systems with the NaCl and CaCl$_{2}$, respectively. Please note that only 350~mM and 450~mM CaCl$_{2}$ concentrations are shown for the Drude and AMOEBA force fields, respectively, while 1000~mM NaCl concentration is shown for all force fields. CHARMM36 NaCl simulations are at 950~mM.}
    \label{fig:ion_density_profiles}
            \todo{SAMULI: We should also show the ion density profiles along membrane normal and compare the ones with 1000 mM NaCl and 350 mM CaCl$_2$
          (concentration in bulk water) to the best profiles in the literature (Fig. 5 in http://dx.doi.org/10.1021/acs.jpcb.7b12510).
          This will show the actual difference in ion binding to the best performing models available.}
          \todo{Batuhan: Data generated from http://dx.doi.org/10.1021/acs.jpcb.7b12510}
\end{figure}

@@

\subsection{Discussion}

\textcolor{red}{For the ion binding discussion: drude calcium overbinds to proteins. this is known here https://pubs.acs.org/doi/full/10.1021/acsphyschemau.1c00039
drude sodium overbinds to proteins, as shown here https://pubs.rsc.org/en/content/articlepdf/2022/ra/d2ra01478e and https://onlinelibrary.wiley.com/doi/abs/10.1002/adts.201800106. Venable also cautioned us against using any ion parameters with the drude force fields. This is important as 1- polarizable force fields are supposed to be better for ion parameters and 2- they are generally used to study the ion channels.}

\subsection{Conclusions}
After comparing the polarizable lipid models against high-resolution experimental data and the best performing non-polarizable force fields, we cannot confirm that adding polarizablity to the force fields has yet solved many of the issues it was hoped to address. Monovalent ion binding is best represented in the non-polarizable CHARMM36 and the polarizable AMOEBA. However, AMOEBA does not model well the divalent ion binding to bilayers. For calcium binding the best models are CHARMM36 and the polarizable CHARMM-Drude2023.
 The electron density profiles of the bilayers are quite well captured by CHARMM-Drude2023 but rather poorly by AMOEBA, within the confines of the approximations used to calculate them from the trajectories. AMOEBA's poor performance regarding electron density curves might be related to the poor representation of lipid tail conformations in the force field, revealed by the order parameters. While CHARMM-Drude2023 captures tail and the headgroup conformations well, the non-polarizable models still outperform when comparing the order parameters. When investigating the conformational dynamics the AMOEBA force field produces results, specifically in the headgroup region of PC lipids, that are in par or better than the best non-polarizable force fields. However, the Drude models exhibit an extremely slow dynamics which leads to deficient sampling and compounding simulation time cost when using the model. In summary, the added complexity of polarizability has not still realised its potential to the fullest; The non-polarizable force fields with longer and more intense refinement behind them still outperform the polarizable models in some respects. Considering the complexity and cost of using these models, polarizability should be parameterized with more thorough consideration to the experimental benchmarks. When doing so, potential for improvement is significant, as exemplified by the refinement of the CHARMM-Drude2017 model to the most recent CHARMM-Drude2023 version.

@@

In the AMOEBA force field, a set of atomic multipoles were first obtained based on the QM electron densities using Stone's distributed multipole analysis~\cite{Stone1981}. The atoms were then assigned polarizabilities according to the AMOEBA atom types and induced dipole contribution was self consistently subtracted from the permanent dipoles~\cite{shi2013proteinamoeba}. The charges (monopoles) were then kept constant while the multipoles were refined against the electrostatic potential energy surface around few conformers of the molecular fragments. %Multipoles are allowed to induce only dipoles at polarizable sites outside their immediate environment (polarization group). 
After electrostatic parameters were set, the Lennard-Jones parameters were taken from previous AMOEBA parameterizations of smaller molecules containing same functional groups~\cite{Ren2011polorganic,shi2011hydration}, and then refined against molecule - water (gas phase) interaction energies. After settling all the nonbonded parameters, dihedral potentials were parameterised based on the QM conformational energy profiles.

@@

Notes on AMOEBA parametrization (this is NOT well explained in the paper)
-bottom up approach
-Stone distributed multipole analysis to get the multipoles from the QM electron densities. Divided into permanent and induced.
-Then iterative fitting to QM-PES for electrostatics
-WdW from 1) optimization to hydration energy calculations and 2) QM fragment-water gas phase calculations.
-dihedrals optimized to QM potentials
-no mention of the screening for polarization? Do they use some AMOEBA value
\end{comment}
\subsubsection{Using a polarizable force field}

The increased level of detail portrayed by explicit treatment of electronic polarizability comes with certain costs. The first cost is the increased computational complexity reduces the simulation speed. For the Drude force field, this slowdown occurs both because the addition of the Drude particles increases the number of interaction pairs and the employed extended dual-Langevin thermostat requires 1~fs integration time step. Current AMOEBA force field can use a multi time-step integration algorithm where the non-electronic interactions are iterated with a 2~fs time step whereas more computationally unstable polarization terms are iterated with a smaller time step. Yet, this usage of a multi-time step integration scheme does not overcome the slowdown resulting from the computationally demanding polarization calculations. As a result, one experiences almost a ten-fold decrease in the simulation speed while using the polarizable force fields. The second cost is due to the limited software support for both running and analyzing the simulations. Currently only OpenMM supports both AMOEBA and Drude force fields. NAMD can only run the Drude force field whereas TINKER is only compatible with AMOEBA force field and does not support semi-isotropic pressure coupling, which decouples the pressure tensor in $xy$ and $z$-direction and is crucial for correctly evaluating the ion binding properties of the bilayers (\todo{This needs a reference, I remember Samuli mentioning this during a chat}). On the other hand, GROMACS, which has been the most popular MD engine within the bilayer simulation field, only has limited support for the Drude polarizable force field via an unofficial git-branch~\cite{drudegithub} but does not support AMOEBA at all.
\begin{comment}
This is particularly important, as GROMACS also comes with many built-in functions for analyzing the trajectories and many of these functions require a GROMACS tpr file and subsequently cannot be utilized for analyzing the trajectories generated by other MD engines. This hinders the researchers from using well-tested analysis tools creating an overhead during the trajectory analysis while making results more prone to error. Therefore, open-source Python libraries for trajectory analysis like MDAnalysis, MDtraj, and LiPhyphilic (based on MDAnalysis) that can work on multiple trajectory and topology formats are gaining more attention.
\end{comment}
Thanks to its recent update, CHARMM-GUI now can generate the topology and input parameters for the Drude force field, which greatly simplifies the employment of this model.
